---
title: "Appendix"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    theme: journal
  word_document:
    keep_md: true
date: "Data Analysis using R programming language"
geometry: "left=1cm,right=1cm,top=1cm,bottom=1cm"
header-includes:
  - \usepackage{booktabs}
  - \usepackage{sectsty} \allsectionsfont{\centering}
  - \DefineVerbatimEnvironment{myVerb}{Verbatim}{numbers=left,numbersep=1mm,frame=lines,framerule=0.4mm,rulecolor=\color{blue}}
  - \usepackage{fontspec}
  - \setmainfont{Arial}
---

```{r setup, include=FALSE}
library(knitr)


knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
opts_chunk$set(comment=NA)
def_hook <- knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  out <- def_hook(x, options)
  return(paste("\\begin{myVerb}\n", x, "\\end{myVerb}", collapse = "\n"))
})
```


\newpage
  
  
  
# Appendix 1  
# RAW DATA WORKSHEET

```{r rawdata, message=FALSE, warning=FALSE}
# Importing necessary libraries
library(tseries)
library(remotes)
library(vars)
library(ggplot2)
library(tidyr)
library(stargazer)
library(psych)
library(reshape2)
library(dplyr)
library(car)
library(forecast)

# Reading the csv file with the values of inflation expectation (one period ahead inflation)
dataset<- read.csv("inf_ir_lead.csv", header = TRUE)

# Converting Date column as  date
dataset$Date <- as.Date(dataset[["Date"]], format = "%d/%m/%Y")

# Printing the entire dataset
knitr::kable(dataset, "latex", booktabs = TRUE, longtable = TRUE, digits = 2)%>%
  kableExtra::kable_styling(font_size = 9)

```

\newpage

# Appendix 2
# DESCRIPTIVE STATISTICS RESULT

```{r descriptive, results='asis', message=FALSE, warning=FALSE}

# Selecting all columns except the date column
selected_columns <- describe(dataset[, 2:ncol(dataset)])

# Computing for the descriptive statistics of variables
summary_stat <- selected_columns[c("mean",
                                   "median",
                                   "sd",
                                   "min",
                                   "max", "range")]

# Printing Summary Statistics
knitr::kable(summary_stat, "latex", booktabs = TRUE, longtable = TRUE, digits = 2)%>%
  kableExtra::kable_styling(font_size = 10)

```


\newpage

# Appendix 3
# LENDING RATES LINE PLOT

```{r ir_line, results='asis', out.width="60%", fig.align='center'}

# Converting dataset into long format 
dataset_long <- dataset[c("Date", c("high", "low"))] %>%
  pivot_longer(cols = -Date, names_to = "Column", values_to = "Value")

# Visualizing the data using line plot
ggplot(dataset_long, aes(x = Date, y = Value, color = Column)) +
  geom_line(linewidth = 0.75) +  
  labs(title = "", x = "Year", y = "Percent per annum") +
  theme_minimal() +
  theme(legend.position = "top", 
        legend.title = element_blank(),
        legend.text = element_text(size = 14)) +
  scale_x_date(date_labels = "%Y", date_breaks  ="4 year") +
  scale_color_manual(values = c("high" = "#336A96", "low" = "#235347"))

```

\newpage

# Appendix 4
# INFLATION EXPECTATIONS LINE PLOTS

```{r all_line3, results='asis', out.width = "60%", fig.align = 'center'}

# Looping through each commodity inflation
for (i in 2:14) {
  
  # Converting dataset into long format
  dataset_long <- dataset[c("Date", names(dataset)[i])] %>%
    pivot_longer(cols = -Date, names_to = "Column", values_to = "Value")
  
  # Get the column name
  col_name <- names(dataset)[i]
  
  # Visualizing the data
  p <- ggplot(dataset_long, aes(x = Date, y = Value, color = col_name)) +
    geom_line(linewidth = 1) +
    labs(title = col_name, x = "Year", y =  "Percent") +
    theme_minimal() +
    theme(legend.position = "none", 
          plot.title = element_text(hjust = 0.5, size = 14)) +
    scale_x_date(date_labels = "%Y", date_breaks = "4 year") +
    scale_color_manual(values = "#336A96")
  
  print(p)
}
```

\newpage

# Appendix 5
# UNIT ROOT TEST AT LEVEL RESULTS

```{r adf, warning=FALSE}

# Looping through each variable
for (i in names(dataset)){
  if (i != "Date"){
    cat("\n-----Unit Root of ",i, "at Level------" )
    
    # Testing for stationarity at level
    print(adf.test(dataset[[i]]))
    
    cat("-----------------------------------------------------------------------------")
  } 
}

```

\newpage

# Appendix 6
# UNIT ROOT TEST AT FIRST DIFFERENCE RESULTS

```{r adf2, warning=FALSE}

# Looping through each variable
for (i in names(dataset)){
  if (i != "Date" ){
    cat("\n-----Unit Root of ",i, " at First Difference------")
    
    # Testing for stationarity at First Difference
    print(adf.test(diff(dataset[[i]])))
    cat("-----------------------------------------------------------------------------")
  } 
}

```

```{r diff, include=FALSE}

# Placeholder dataframe to be replaced by first difference
# Exclude first column because first difference
dataset_diff <- dataset[2:314,]

# Looping though each column
for (i in names(dataset_diff)){
  if (i != "Date"){
    
    # To take the first difference 
    data = diff(dataset[[i]])
    
    # And replace data in the placeholder dataframe
    dataset_diff[[i]] = data
  } }

```

\newpage

# Appendix 7
# AKAIKE INFORMATION CRITERION RESULTS

```{r akaike}

# Looping through the first difference dataset
for (i in names(dataset_diff)){
  if (i != "Date" && i != "high"  && i != "low"  ){
    
    cat("\n-----High lending rate and ", i, "-----\n")
    
    # To get the optimal lag length of every combination of high lending rate and commodity inflation
    print(VARselect(data.frame(dataset_diff[["high"]], dataset_diff[[i]])))
    
    cat("-----------------------------------------------------------------------------")
    
    cat("\n-----Low lending rate and ", i, "------\n")
    
    # To get the optimal lag length of every combination of low lending rate and commodity inflation
    print(VARselect(data.frame(dataset_diff[["low"]], dataset_diff[[i]])))
    
    cat("-----------------------------------------------------------------------------")
  } 
}

```

\newpage

# Appendix 8
# GRANGER CAUSALITY TEST RESULTS

```{r granger1}
# Is lending interest rates granger cause commodity inflation expectation?
# The lag order is based on AIC
grangertest( dataset_diff[['all']] ~ dataset_diff[['high']], order=3)
grangertest( dataset_diff[['all']] ~ dataset_diff[['low']], order=3)
grangertest( dataset_diff[['food']] ~ dataset_diff[['high']], order=1)
grangertest( dataset_diff[['food']]~ dataset_diff[['low']], order=4)
grangertest( dataset_diff[['alco']] ~ dataset_diff[['high']], order=3)
grangertest( dataset_diff[['alco']]~ dataset_diff[['low']], order=3)
grangertest( dataset_diff[['cloth']] ~ dataset_diff[['high']], order=9)
grangertest( dataset_diff[['cloth']]~ dataset_diff[['low']], order=4)
grangertest( dataset_diff[['house']] ~ dataset_diff[['high']], order=1)
grangertest( dataset_diff[['house']]~ dataset_diff[['low']], order=2)
grangertest( dataset_diff[['furnish']] ~ dataset_diff[['high']], order=5)
grangertest( dataset_diff[['furnish']]~ dataset_diff[['low']], order=4)
grangertest( dataset_diff[['health']] ~ dataset_diff[['high']], order=7)
grangertest( dataset_diff[['health']]~ dataset_diff[['low']], order=10)
grangertest( dataset_diff[['transpo']] ~ dataset_diff[['high']], order=2)
grangertest( dataset_diff[['transpo']] ~ dataset_diff[['low']], order=5)
grangertest( dataset_diff[['ict']] ~ dataset_diff[['high']], order=2)
grangertest( dataset_diff[['ict']]~ dataset_diff[['low']], order=10)
grangertest( dataset_diff[['rec']] ~ dataset_diff[['high']], order=7)
grangertest( dataset_diff[['rec']] ~ dataset_diff[['low']], order=7)
grangertest( dataset_diff[['educ']] ~ dataset_diff[['high']], order=7)
grangertest( dataset_diff[['educ']] ~ dataset_diff[['low']],order=7)
grangertest( dataset_diff[['restau']] ~ dataset_diff[['high']], order=4)
grangertest( dataset_diff[['restau']] ~ dataset_diff[['low']], order=4)
grangertest( dataset_diff[['personal']] ~ dataset_diff[['high']], order=10)
grangertest( dataset_diff[['personal']] ~ dataset_diff[['low']], order=10)

```


\newpage

# Appendix 9
# R PROGRAMMING LANGUAGE ALGORITHM FOR FORECASTING INFLATION EXPECTATIONS WITH HIGH LENDING RATE

```{r high_AR, fig.align="center", warning=TRUE, results="asis", out.width="75%"}

# Loop over each significant variables in Granger Causality Test
high_variables <- c("all", "cloth", "health", "transpo", "ict", "rec", "personal")
fullname <- c("ALL ITEMS", "CLOTHING AND FOOTWEAR", "HEALTH", "TRANSPORT", "INFORMATION AND COMMUNICATION", "RECREATION, SPORT AND CULTURE", "PERSONAL CARE, AND MISCELLANEOUS GOODS AND SERVICES")


# Appendix count for the title
Appendix = 10

for (inf_exp in high_variables) {
  
  # Create a monthly time series object for the variables
  ts_inf_exp <- ts(dataset[1:290,inf_exp],
                   start = c(1997, 9), 
                   end = c(2021, 10),
                   frequency = 12)
  ts_high <- ts(dataset[1:290,"high"],
                start = c(1997, 9),
                end = c(2021, 10),
                frequency = 12)
  
  # Create a monthly time series object for data visualization 
  actual_inf_exp <- ts(dataset[289:314,inf_exp],
                       start = c(2021, 11),
                       end = c(2023, 10),
                       frequency = 12)
  actual_inf_exp_long <- ts(dataset[1:290,inf_exp],
                            start = c(1997, 10),
                            end = c(2021, 10),
                            frequency = 12)
  
  # Create a monthly time series object for testing forecasting
  actual_high <- ts(dataset[291:314,"high"],
                    start = c(2021, 11),
                    end = c(2023, 10),
                    frequency = 12)
  
  cat("\\newpage")
  
  cat("# Appendix",Appendix, "\n")
  cat("# FORECASTING ", fullname[Appendix - 9], "INFLATION EXPECTATION WITH HIGH LENDING RATE\n")
  Appendix = Appendix + 1
  
  cat(" HIGH -> ", toupper(inf_exp)," Inflation Expectation\n") 
  
  # Linear Regression with Autoregressive term fit
  ar_model <- arima(ts_inf_exp,
                    order = c(1, 0, 0),
                    xreg = ts_high)
  
  # Extract residuals from the fitted AR model
  ar_residuals <- ts(residuals(ar_model),
                     start = c(1997, 9),
                     end = c(2021, 10), 
                     frequency = 12)

  cat("\n-------- RESIDUALS TIME PLOT ----------" ) 
  
  # Setting up the plot
  plot(ar_residuals,
       ylab = "Residuals",
       xlab = "Year-Month",
       xaxt = "n",
       main = paste("HIGH -> ", toupper(inf_exp)," Inflation Expectation"))
  
  dates <- as.Date(time(ar_residuals))
  axis_dates <- time(ar_residuals)[seq(1, length(ar_residuals), by = 4)]
  axis_labels <- format(dates[seq(1, length(ar_residuals), by = 4)], "%Y-%m")

  axis(1, at = axis_dates,labels = axis_labels,las = 2, cex.axis = 0.6)

  cat("\n\n-------- HETEROCEDASTICITY TEST ----------\n\n" )
  print(bptest(lm(dataset[2:290,inf_exp] ~ dataset[2:290,"high"] + dataset[1:289,inf_exp])))
  cat("\n")
  
  cat("\n------------- NORMALITY TEST ----------------\n " ) 
  # Normality
  print(jarque.bera.test(ar_residuals))
  cat("\n")
  
  cat("\n -------- AUTOCORRELATION TEST ----------\n " ) 
  # Serial Autocorrelation
  print(Box.test(ar_residuals, lag = 10, type = "Ljung-Box"))
  cat("\n")
  cat("\n -------- FORECASTING SUMMARY  ---------- \n " ) 
  print(summary(ar_model))
  cat("\n")
  
  # Forecast the inflation expectation from Nov 2021 to Oct 2025
  inf_forecast <- predict(ar_model, newxreg=actual_high, n.ahead=24)
  inf_forecast_values <- inf_forecast$pred
  inf_forecast_se <- inf_forecast$se
  
  # Extracting dates
  dates <- as.Date(time(inf_forecast_values))
  
  # Plotting
  plot(inf_forecast_values, 
       main = paste("2-year Forecast for",toupper(inf_exp),"Inflation Expectation"), 
       xlab = "Year-Month",
       ylab = "Inflation Expectation",
       xaxt = "n",col = "red",
       lty = 1) 
  
      # Customize x-axis labels with year-month format
      axis(1,
           at = time(inf_forecast_values)[seq(1, length(inf_forecast_values), by = 4)],
           labels = format(dates[seq(1, length(inf_forecast_values), by = 4)], "%Y-%m"))

  # Add thin gray lines matching the y-axis labels
  y_ticks <- axTicks(2)
  for (y in y_ticks) {
    abline(h = y, col = "gray", lty = 2, lwd = 0.2)}

  # Add value labels every 3 data points
  every_third <- seq(1, length(inf_forecast_values), by = 3)
  text(x = time(inf_forecast_values)[every_third],
       y = inf_forecast_values[every_third],
       labels = round(inf_forecast_values[every_third], 2),
       pos = 3, col = "black", cex = 0.8)
  
  # Variables will be used to zoom out the graph
  overall_max <- max(max(coredata(inf_forecast_values)), 
                     max(actual_inf_exp_long), max(inf_forecast_values - 1.96 * inf_forecast_se), 
                     max(inf_forecast_values + 1.96 * inf_forecast_se))
  overall_min <- min(min(coredata(inf_forecast_values)), min(actual_inf_exp_long), 
                     min(inf_forecast_values - 1.96 * inf_forecast_se), 
                     min(inf_forecast_values + 1.96 * inf_forecast_se))
  
  # Plotting the actual values of inflation expectation 
  plot.ts(actual_inf_exp_long,
          main = paste("Forecast for", toupper(inf_exp), "Inflation Expectation"),
          xlab = "Year",
          ylab = "Inflation Expectation",
          lty = 1,
          ylim = c(overall_min - 0.5, overall_max + 0.5))
  lines(actual_inf_exp, col = "black", lty = 3)
  
  # Add the forecast on the end and confidence intervals 
  lines(inf_forecast_values, col = "red", lwd = 2) 
  lines(inf_forecast_values - 1.96 * inf_forecast_se, col = 4, lty = 1, lwd = 2)
  lines(inf_forecast_values + 1.96 * inf_forecast_se, col = 4, lty = 1, lwd = 2)
  
  # Add thin gray lines matching the y-axis labels
  y_ticks <- axTicks(2)
  for (y in y_ticks) {
    abline(h = y, col = "gray", lty = 2, lwd = 0.2)
  }

  # Add legend
  legend("top",
         legend = c("Actual Values", "Forecast", "Confidence Interval"),
         col = c("black", "red", "blue"),
         lty = c(1, 1, 1),
         lwd = c(2, 2, 1),
         cex = 0.6, horiz = TRUE, 
         bg = "transparent") 
  }
```

\newpage

# Appendix 17
# R PROGRAMMING LANGUAGE ALGORITHM FOR FORECASTING INFLATION EXPECTATIONS WITH LOW LENDING RATE

```{r low_AR, fig.align="center", warning=TRUE, results="asis", out.width="75%"}

# Loop over each significant variables in Granger Causality Test
low_variables <- c("all", "cloth", "health", "transpo", "ict", "rec", "personal", "food", "educ")

fullname <- c("ALL ITEMS", 
              "CLOTHING AND FOOTWEAR",
              "HEALTH", "TRANSPORT",
              "INFORMATION AND COMMUNICATION",
              "RECREATION, SPORT AND CULTURE",
              "PERSONAL CARE, AND MISCELLANEOUS GOODS AND SERVICES",
              "FOOD AND NON-ALCOHOLIC BEVERAGES",
              "EDUCATION SERVICES")

# Appendix count for the title
Appendix = 18

for (inf_exp in low_variables) {
  
  # Create a monthly time series object for the variables
  ts_inf_exp <- ts(dataset[1:290,inf_exp],
                   start = c(1997, 9), 
                   end = c(2021, 10),
                   frequency = 12)
  ts_low <- ts(dataset[1:290,"low"],
                start = c(1997, 9),
                end = c(2021, 10),
                frequency = 12)
  
  # Create a monthly time series object for data visualization 
  actual_inf_exp <- ts(dataset[289:314,inf_exp],
                       start = c(2021, 11),
                       end = c(2023, 10),
                       frequency = 12)
  actual_inf_exp_long <- ts(dataset[1:290,inf_exp],
                            start = c(1997, 10),
                            end = c(2021, 10),
                            frequency = 12)
  
  # Create a monthly time series object for testing forecasting
  actual_low <- ts(dataset[291:314,"low"],
                    start = c(2021, 11),
                    end = c(2023, 10),
                    frequency = 12)
  
  cat("\\newpage")
  cat("# Appendix",Appendix,"\n")
  cat("# FORECASTING ", fullname[Appendix - 17], "INFLATION EXPECTATION WITH LOW LENDING RATE\n")
  Appendix = Appendix + 1
  cat(" low -> ", toupper(inf_exp)," Inflation Expectation\n") 
  
  # Linear Regression with Autoregressive term fit
  ar_model <- arima(ts_inf_exp,
                    order = c(1, 0, 0),
                    xreg = ts_low)
  
  # Extract residuals from the fitted ARIMA model
  ar_residuals <- ts(residuals(ar_model),
                     start = c(1997, 9),
                     end = c(2021, 10), 
                     frequency = 12)
  
  cat("\n-------- RESIDUALS TIME PLOT ----------" ) 
  
  # Setting up the plot
  plot(ar_residuals,
       ylab = "Residuals",
       xlab = "Year-Month",
       xaxt = "n",
       main = paste("LOW -> ", toupper(inf_exp)," Inflation Expectation"))
  
  dates <- as.Date(time(ar_residuals))
  axis_dates <- time(ar_residuals)[seq(1, length(ar_residuals), by = 4)]
  axis_labels <- format(dates[seq(1, length(ar_residuals), by = 4)], "%Y-%m")
  
  axis(1, at = axis_dates,labels = axis_labels,las = 2, cex.axis = 0.6)
  
  cat("\n\n-------- HETEROCEDASTICITY TEST ----------\n\n" )
  print(bptest(lm(dataset[2:290,inf_exp] ~ dataset[2:290,"low"] + dataset[1:289,inf_exp])))
  cat("\n")
  
  cat("\n------------- NORMALITY TEST ----------------\n " ) 
  # Normality
  print(jarque.bera.test(ar_residuals))
  cat("\n")
  
  cat("\n -------- AUTOCORRELATION TEST ----------\n " ) 
  # Serial Autocorrelation
  print(Box.test(ar_residuals, lag = 10, type = "Ljung-Box"))
  cat("\n")
  cat("\n -------- FORECASTING SUMMARY  ---------- \n " ) 
  print(summary(ar_model))
  cat("\n")
  
  # Forecast the inflation expectation from Nov 2021 to Oct 2025
  inf_forecast <- predict(ar_model, newxreg=actual_low, n.ahead=24)
  inf_forecast_values <- inf_forecast$pred
  inf_forecast_se <- inf_forecast$se
  
  # Extracting dates
  dates <- as.Date(time(inf_forecast_values))
  
  # Plotting
  plot(inf_forecast_values, 
       main = paste("2-year Forecast for",toupper(inf_exp),"Inflation Expectation"), 
       xlab = "Year-Month",
       ylab = "Inflation Expectation",
       xaxt = "n",col = "red",
       lty = 1) 
  
  # Customize x-axis labels with year-month format
  axis(1,
       at = time(inf_forecast_values)[seq(1, length(inf_forecast_values), by = 4)],
       labels = format(dates[seq(1, length(inf_forecast_values), by = 4)], "%Y-%m"))
  
  # Add thin gray lines matching the y-axis labels
  y_ticks <- axTicks(2)
  for (y in y_ticks) {
    abline(h = y, col = "gray", lty = 2, lwd = 0.2)}
  
  # Add value labels every 3 data points
  every_third <- seq(1, length(inf_forecast_values), by = 3)
  text(x = time(inf_forecast_values)[every_third],
       y = inf_forecast_values[every_third],
       labels = round(inf_forecast_values[every_third], 2),
       pos = 3, col = "black", cex = 0.8)
  
  # Variables will be used to zoom out the graph
  overall_max <- max(max(coredata(inf_forecast_values)), 
                     max(actual_inf_exp_long), max(inf_forecast_values - 1.96 * inf_forecast_se), 
                     max(inf_forecast_values + 1.96 * inf_forecast_se))
  overall_min <- min(min(coredata(inf_forecast_values)), min(actual_inf_exp_long), 
                     min(inf_forecast_values - 1.96 * inf_forecast_se), 
                     min(inf_forecast_values + 1.96 * inf_forecast_se))
  
  # Plotting the actual values of inflation expectation 
  plot.ts(actual_inf_exp_long,
          main = paste("Forecast for", toupper(inf_exp), "Inflation Expectation"),
          xlab = "Year",
          ylab = "Inflation Expectation",
          lty = 1,
          ylim = c(overall_min - 0.5, overall_max + 0.5))
  lines(actual_inf_exp, col = "black", lty = 3)
  
  # Add the forecast on the end and confidence intervals 
  lines(inf_forecast_values, col = "red", lwd = 2) 
  lines(inf_forecast_values - 1.96 * inf_forecast_se, col = 4, lty = 1, lwd = 2)
  lines(inf_forecast_values + 1.96 * inf_forecast_se, col = 4, lty = 1, lwd = 2)
  
  # Add thin gray lines matching the y-axis labels
  y_ticks <- axTicks(2)
  for (y in y_ticks) {
    abline(h = y, col = "gray", lty = 2, lwd = 0.2)
  }
  
  # Add legend
  legend("top",
         legend = c("Actual Values", "Forecast", "Confidence Interval"),
         col = c("black", "red", "blue"),
         lty = c(1, 1, 1),
         lwd = c(2, 2, 1),
         cex = 0.6, horiz = TRUE, 
         bg = "transparent") 
}
```


